---
output:
  word_document: default
  html_document: default
---


# Installing the required libraries:

```{r}
#install.packages("lubridate")
#install.packages("broom")
#install.packages("survival")
#install.packages("survminer")

library(tidyverse)
library(lubridate)
library(broom)
library(survival)
library(survminer)

```


# Data Import:
```{r}
    setwd("C:/Users/Samer/Desktop/Survival_Analysis")
   data <- read.csv('DSTI_survey.csv',header = TRUE)

   # Data Exploraration
   head(data)
   str(data)
   typeof(data)

```
# Data Cleaning and Preparation


```{r}
# Rename variables 
names(data)[1]  = "srv_date"      #TimeStamp
names(data)[2]  = "yob"           #Year of Birth
names(data)[3]  = "smkr"          #Were you ever a smoker? 
names(data)[4]  = "y_str_smk"     #Year when first started smoking
names(data)[5]  = "y_stp_smk"     #Year when stopped smoking
names(data)[6]  = "date_str_srch" #When did you start looking for an internship
names(data)[7]  = "sex"           #Sex
names(data)[8]  = "date_stp_srch" #When did you stopped looking for an internship
names(data)[9]  = "fnd_intr"      #Have you found an internship?
names(data)[10] = "edu"           #Education: background (pick a main one you identify with)
names(data)[11] = "yoe"           #Years of education
names(data)[12] = "children"      #Do you have children?
names(data)[13] = "cohort"        #Cohort  

head(data)
```


# Converting data to the correct format:

```{r}

# Convert Year to Date 
data[,2]           = year(ymd(data[,2], truncated = 2L))
data[,1]           = as_date(as.POSIXct(data[,1], format='%m/%d/%Y %H:%M'))
data$y_str_smk     = year(ymd(data$y_str_smk, truncated = 2L))
data$y_stp_smk     = year(ymd(data$y_stp_smk, truncated = 2L))
data$date_str_srch = as_date(as.POSIXct(data$date_str_srch, format='%m/%d/%Y'))
data$date_stp_srch = as_date(as.POSIXct(data$date_stp_srch, format='%m/%d/%Y'))

# Convert Categorical Variables into Factors
data$sex       = as.factor(data$sex) 
#data$fnd_intr  = as.factor(data$fnd_intr) 
data$children  = as.factor(data$children) 
data$cohort    = as.factor(data$cohort) 

# Turning smkr variable to binary with (yes,no)
data$smkr= ifelse(data$smkr != "No", "Yes", data$smkr)
data$smkr      = as.factor(data$smkr) 

data$fnd_intr= ifelse(data$fnd_intr == "Yes", 1, 0)


#Sanity Checks
str(data)
head(data)
tail(data)
data
```


```{r}
# Cleaning up Education Variable
data$edu= ifelse(data$edu == "Mathematics, Physics, Chemistry, Computer Science, Statistics", "math", data$edu)
data$edu= ifelse(data$edu == "Medicine, Biology", "bio"                                             , data$edu)
data$edu= ifelse(data$edu == "Literature, History, Philosophy", "lit"                               , data$edu)
data$edu= ifelse(data$edu == "Finance, Economy", "fin"                                              , data$edu)
data$edu= ifelse(data$edu == "Business, Management", "mgm"                                          , data$edu)
data$edu= ifelse(data$edu == "Other", "oth"                                                         , data$edu)

# Converting Education variable into factor
data$edu= as.factor(data$edu) 

#Sanity Checks
head(data)
tail(data)

```


# Creating censored Duration Variable 
 As we want to analyze the time to internship
 first we need the duration variable which will be the timestamp minus date when started search,
 Noting that if the stop searching date is more that timestamp, we replaced it with the timestamp date instead.

```{r}
data$date_stp_srch= as.Date(ifelse(data$date_stp_srch > data$srv_date, data$srv_date, data$date_stp_srch), origin = "1970-01-01")
```

```{r}

# Creating Search Duration Variable

data$srch_dur=
ifelse(data$fnd_intr == "1",
 difftime(data$date_stp_srch, data$date_str_srch, units = "days"),
 ifelse(is.na(data$date_stp_srch),
 difftime(data$srv_date, data$date_str_srch, units = "days"),
 difftime(data$date_stp_srch, data$date_str_srch, units = "days")))

data$srch_dur

```

# Elemenation of obervations are not suitable to enter our model

1. Remove students with no Start Date 
```{r}
# Removing Students with no Start Date
data<-data[!(is.na(data$date_str_srch)), ]
data


```


2. remove Students with Zero duration and found an Internship 

```{r}
#Students with Zero duration and found an Internship
data<-data[!(data$srch_dur=="0" & data$fnd_intr=="1"),]
data
```

3. Remove students with start date after Time stamp date, and this is based on their assupmtion not on reality,
and this considered as Right cesoring which is not out of the scope of our analysis.

```{r}
# Remove people with Internship Yes and stop search date NA
data<-data[!(data$date_str_srch > data$srv_date),]
data
```


```{r}

summary(data$srch_dur)
hist(data$srch_dur)
boxplot(data$srch_dur)
```
we can see some outliers here,
the first try is to remove any observation more that 600 days of search duration

```{r}
# Remove observations with seaerch duration >500
data<-data[!(data$srch_dur >= 500),]
boxplot(data$srch_dur)
data
```


# Q1. How many students partecipated in the interview?
 84 students 

# Q2. After data preparation, how many samples are usable for data analysis? How many samples were dropped (if  any), and why?

Raw Dataset included 84 rows, after processing, some of variables were dropped due to the following reasons:

1. Students who don't have start search date (18 students)
2. Students with Zero duration and found an Internship (1 student)
3. students with start date after Time stamp date, and this is based on their assumption not on reality,
and this considered as Right censoring which is not out of the scope of our analysis.(7 students)
4. 2 outliers removed, search duration >500 days (2 students)

The final number of students of samples usable for analysis is 54



# Survival Function 

Now since we defined our Censoring variable (fnd_intr) and Duration variable (srch_dur)
we can start with the Survival Function

```{r}
# Surviaval Material
srv_mat <- with(data, Surv(srch_dur ,fnd_intr))

 KM <- survfit(srv_mat ~ 1, data = data)
 KM
 summary(KM)
 
 plot(KM, 
     # fun = "F",
      main = "Kaplan-Meier estimator",
      xlab = "Internship Search Duration (Days)", 
      ylab = "Cumulated incidence of Students")
```


#FROM: cLASS 3 - 12:20

#Q3.How long does it take to obtain an internship?
#   Please report the median time (with a confidence interval), 
#   total number of students at the baseline, 
#   the total number of events observed, and the total number of censored observations.
   
1. Median is 92 days, 3 months approximately.
with Confidence Interval of [61, 305]

2. Total number of students at the baseline ?
54 students

3. Total Number of Events Obsereved = 23
4. Total Number of Events Censored = 31
  
  
  
 
# Q4. Of these variables,. which ones have the most impact on the time to obtain an internship, and in which # direction: 
# 1.cohort, 
# 2.age, 
# 3.educational background, 
# 4.having or not having children.
 
 
 
# 1. Analyzing by Cohort groups


```{r}
# Log Rank test - Cohort
surv_cohort_LR <- survdiff(Surv(srch_dur ,fnd_intr) ~ cohort, data = data)
surv_cohort_LR
```
Cohort is Significant (very small p-value)



# 2. Analyzing by Age  groups

We have to create the age variable, which will be the the timestamp (srv_date) minus year of birth
then we will create age groups 


```{r}
#Calculating Age

data$age = year(data$srv_date) - data$yob
data$age
summary(data$age)
```
Creating age groups:

```{r}
# Creating Age groups 

data$age_gr= ifelse(between(data$age,20,29), "20-29",
                    ifelse(between(data$age, 30, 39), "30-39",
                           ifelse(between(data$age, 40, 49), "40-49",
                                  ifelse(between(data$age, 50, 59), "50-59",
                                                ifelse(data$age >59 , ">=60", "Error")))))
data$age_gr

```
# Age Impact
```{r}
# Log Rank test - age
surv_age_LR <- survdiff(Surv(srch_dur ,fnd_intr) ~ age_gr, data = data)
surv_age_LR

```


# 3. Education Impact :

```{r}
# Log Rank test - Education
surv_edu_LR <- survdiff(Surv(srch_dur ,fnd_intr) ~ edu, data = data)
surv_edu_LR

```


# 3. Analyzing Children groups:

```{r}
# Log Rank test - Children
surv_children_LR <- survdiff(Surv(srch_dur ,fnd_intr) ~ children, data = data)
surv_children_LR

```


After analyzing the impact of (cohort, age, children, edcation background)
it found that there is only Chorot is significent which p-value is <0.05



# Mesuring Impact of the 4 Variables using coxph regression Model:

```{r}

coxph <- coxph(Surv(srch_dur ,fnd_intr) ~ cohort + age + edu + children, data = data)

coxph
summary(coxph)
```

The only significant variables are S20, S19, A20 according to the model, all other variables are not significent.



# Bounus Question: Can you build a predictive model to identify students at high risk of a long search? How well does your model perform?"

As we saw earlier, there are no enough relible signifecnt variables we could use to build a predective model, as cohort variable alone is not enough to explain or to pridect the time to internship.

So, based on these factors and current data situation, it's not possible to build a predctive model.







